\chapter{Discussion and Conclusion}
\label{chp:conclusion}

\input{chapters/conclusion/discussion.tex}
% \input{chapters/conclusion/future.tex}

\section{Conclusion}
\label{sec:conclusion}

In this thesis, we advanced principled, automated methods for smart contracts by exploiting their structural properties and transaction-level interactions. At the code level, we introduced algebro-geometric techniques to synthesize tight polynomial gas upper bounds with soundness (and semi-completeness) that outperformed prior work in both tightness and applicability. We further strengthened compiler-level gas superoptimization by integrating a dynamic-programming enhancement into the state of the art, more than doubling baseline gas savings on widely used contracts. At the interaction level, we designed revenue-maximizing block-construction algorithms: for Cardano, an exact parameterized approach that runs efficiently on sparse transaction graphs and achieves a 55.68\% improvement; and for Ethereum, a randomized framework that increases miner revenue by an average of 73.45\% per block. For decentralized exchange rates, we leveraged the tree-like topology of liquidity pools, enabling support for instantaneous trade sizes and delivering up to four orders of magnitude performance improvements over prior work. Beyond engineering, we analyzed Bitcoin's security under derivatives, showing that block-reversion attacks can be economically enabled with far less than majority hash power.
Taken together, these contributions show that exploiting the unique structure of smart contracts can yield techniques that are simultaneously grounded in formal principles and effective at scale. At the same time, the thesis opens several directions for further investigation. In particular, the compiler-level optimization results should be viewed as an initial step and can be substantially improved by applying the same foundational principles. More broadly, a range of smart-contract analysis tasks remain underexplored, and the tools developed here appear well-suited to be adapted.
