% Supervisors
I was privileged to have Amir Goharshady as my PhD supervisor. His insights and guidance were instrumental in my growth as a researcher. He consistently supported me in navigating career choices, grant applications, and job opportunities.
Amir’s dedication to his students goes far beyond standard expectations. He truly embodies the revered title of ``Ostad'' in the classical academic tradition.
Our long discussions about the right and wrong paths, whether in research, career, or life itself, have always been invaluable to me.

I am grateful to my co-supervisor, Jiasi Shen, for her consistent support throughout my time at HKUST. Thanks are also due to S. Akshay and Supratik Chakraborty for hosting my visit to IIT Bombay and for sharing valuable perspectives on both research and career development. I appreciate the advice of \DJ or\dj e \v{Z}ikeli\'{c} regarding my professional development. Finally, I would like to thank Dimitris Papadopoulos, Guillermo Perez, He Junxian, Ke Yi, Maximilian Nitzschner, and Charles Zhang for agreeing to chair my thesis defense sessions.

% COllaborators
During my PhD, I had the privilege of working alongside exceptionally talented people. I am forever in debt for what I have learned from Sergei Novozhilov, Harshit Motwani, and Togzhan Barakbayeva. We navigated countless challenges together, pushing through late nights to meet tight deadlines, sharing work lunches, and coming up with bold ideas in our stormy meetings. The conversations and the work we shared deeply influenced many chapters of this thesis.

% Alpacas
I am also proud to be part of the ALPACAS group and appreciate the support from my colleagues: Jonas Ballweg, Xuran Cai, Zhuo Cai, Giovanna Kobus Conrado, Pavel Hudec, S. Hitarth, Kerim Kochekov, Chun Kit Lam, Pingjiang Li, Richard Magambo, Zhaorun Lin, Tian Shu, and Ahmed Zaher.

% Family
I owe a profound debt of gratitude to my parents, Marzieh Javameh and Saeid Farokhnia, for their unwavering support throughout my life. My brother Sepher likewise deserves recognition for his support\footnote{Though I must note that fewer hours spent debating about Middle-earth could have helped me write a couple more papers!}.
% New Family
I must also express my deepest appreciation to my fianc\'ee, Sanaz Safaei. This doctorate would not have been possible without her kind support and encouragement. It is only fitting that our next chapter begins as this one closes.

% Teachers
I am grateful to my teachers over the years, whose guidance has brought me to this point. Mahdi Kazemi and Hossein Boomari introduced me to the world of computer science.
Mohammadreza Shabanali’s writings have been a constant source of direction for me, fundamentally shaping how I think, learn, and pursue personal growth.
In the early stages of my career, I benefited greatly from the advice of Behnam Shakibafar and Reza Sajjadi.

Finally, I am grateful to HKUST and the Ethereum Foundation, which financially supported my PhD, and to the Alexander von Humboldt Foundation, which awarded me a research fellowship.