The following is a list of all publications from the PhD period. Following the norms of theoretical computer science, authors are listed in alphabetical order.

\begin{enumerate}
    \item T. Barakbayeva, S. Farokhnia, A.K. Goharshady, S. Novozhilov. \textit{Boosting Gas Revenues of Ethereum Miners}. IEEE/ACM International Conference on Software Engineering, ICSE 2026.
    
    \item S. Farokhnia, S. Novozhilov, S. Safaei, J. Shen. \textit{Hermes: Scalable and Robust Structure-Aware Optimal Routing for Decentralized Exchanges}. IEEE International Conference on Blockchain, Blockchain 2025.
    
    \item S. Akshay, S. Chakraborty, S. Farokhnia, A.K. Goharshady, H.J. Motwani, \DJ. \v{Z}ikeli\'{c}. \textit{LP-Based Weighted Model Integration over Non-Linear Real Arithmetic}. International Joint Conference on Artificial Intelligence, IJCAI 2025.
    
    \item T. Barakbayeva, S. Farokhnia, A.K. Goharshady, P. Li, Z. Lin. \textit{Improved Gas Optimization of Smart Contracts}. International Conference on Fundamentals of Software Engineering, FSEN 2025.
    
    \item T. Barakbayeva, S. Farokhnia, A.K. Goharshady, M. Gufler, S. Novozhilov. \textit{Pixiu: Optimal Block Production Revenues on Cardano}. IEEE International Conference on Blockchain, Blockchain 2024.
    
    \item S. Farokhnia, A.K. Goharshady. \textit{Options and Futures Imperil Bitcoin’s Security}. IEEE International Conference on Blockchain, Blockchain 2024.
    
    \item Z. Cai, S. Farokhnia, A.K. Goharshady, S. Hitarth. \textit{Asparagus: Automated Synthesis of Parametric Gas Upper-bounds for Smart Contracts}. ACM Conference on Object-Oriented Programming, Systems, Languages, and Applications, OOPSLA 2023.
\end{enumerate}